Composing Robustly Safe Compilers

### 2023-06-11 popl/main.tex
Upto section 3

\usepackage[colorinlistoftodos]{todonotes}
\usepackage{cleveref}
\usepackage{glossaries}
\usepackage{listings}
\usepackage{xspace}

@@

\section{Background}\label{sec:background}
We give a brief introduction to trace properties, monitors, and property satisfaction (\Cref{subsec:bg:tprop}).
With property satisfaction, we re-introduce the concept of robust satisfaction and its compiler-driven preservation (\Cref{subsec:bg:rtp}).

\subsection{Trace-Based Properties}\label{subsec:bg:tprop}
\subsection{Robust Trace-Property Preservation}\label{subsec:bg:rtp}

\section{Composing Properties}\label{sec:compprop}
\section{Composing Robustly Secure Compilers}\label{sec:compcomp}
This section studies different kinds of secure compiler compositions and their corresponding proofs.
Most notably, it looks at the sequential composition of secure compilers and shows a theorem that allows for modular proof engineering: compilers need not to robustly preserve the same property in order to be securely composed.
Moreover, it demonstrates that the order does not matter for composing compilers with the same source and target language.

>>>>

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage{cleveref}
\usepackage{glossaries}
\usepackage{listings}
\usepackage{xspace}

@@

\section{Background}\label{sec:background}
\Cref{subsec:bg:tprop} introduces trace properties, monitors, and property satisfaction.
\Cref{subsec:bg:rtp} extends the property satisfaction with robustness and shows its compiler-driven preservation, as presented in~\cite{abate2019jour}.

\subsection{Trace-Based Properties, Programming Languages, and Property Satisfaction}\label{subsec:bg:tprop}

\emph{Events} $\event$ are an abstract representation of observable actions a program can do during its execution.
Typically, one always has an empty observation $\emptyevent$ used to represent unimportant actions.
A sequence of events $\trace$ is a \emph{trace}.
So, throughout the paper, we use $\overline{\cdot}$ to mark meta-variables as sequences.
Empty sequences are denoted with $\hole{\cdot}$ and concatenation of traces $\trace[_{1}],\trace[_{2}]$ is written as $\trace[_{1}]\cdot\trace[_{2}]$.
For sake of brevity, we use this notation also for pre- and appending events to traces.

A trace-based property $\pi$ is a set of admissable traces.
For a trace to be admissable, a predicate evaluates to true whenever $\trace\in\pi$.
Classes of properties $\class$, commonly referred to as hyperproperties in literature, are sets of properties.
The \emph{lifting} of a property $\pi$ is simply its powerset and denoted as $\lift{\pi}$.
To define property satisfaction, we first look at programming languages.

For languages used in this paper, we consider a small-step style semantics.
That is, there is a step relation between states $\statevar$ and an event $\event$: $\stepto[]{\statevar[_{1}]}{\statevar[_{2}]}{\event}$.
Also, a program's syntax $p$ is encoded in the state $\statevar$ which we write as $p\in\statevar$.
For states $\statevar$, we assume there exists a judgement $\isval{\statevar}$ that determines wether $\statevar$ is a final state or not.
\Cref{fig:starstep:and:nstep}\MK{appendix?} defines the star-step relation $\stepto{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$, essentially modelling a big-step style evaluation, and the $n$-step relation $\stepton{\statevar[_{1}]}{\statevar[_{2}]}{\trace}$ that evaluates a program for $n$ steps.
Both relations filter $\varepsilon$ events.

\begin{figure}[t]
  \judgbox{\stepto{r}{r'}{\trace}}{,,State $\statevar$ steps to final state $\statevar[']$, emitting events $\trace$.''}$\;$\\
  %
  \typerule{star-step-refl}{
    \isval{\statevar}
  }{
    \stepto{\statevar}{\statevar}{\hole{\cdot}}
  }{star-step-refl}
  %
  \typerule{star-step-trans}{
    \event\not=\varepsilon &
    \stepto[]{\statevar}{\statevar[']}{\event} &
    \stepto{\statevar[']}{\statevar['']}{\trace}
  }{
    \stepto{\statevar}{\statevar['']}{\event\cdot\trace}
  }{star-step-trans}
  %
  \typerule{star-step-ign}{
    \stepto[]{\statevar}{\statevar[']}{\varepsilon} &
    \stepto{\statevar[']}{\statevar['']}{\trace}
  }{
    \stepto{\statevar}{\statevar['']}{\trace}
  }{star-step-ign}

  \judgbox{\stepton{\statevar}{\statevar[']}{\trace}}{,,State $r$ does $n$ steps to final state $\statevar[']$, emitting events $\trace$.''}$\;$\\
  %
  \typerule{n-step-refl}{
  }{
    \stepton[0]{\statevar}{\statevar}{\hole{\cdot}}
  }{n-step-refl}
  %
  \typerule{n-step-trans}{
    \event\not=\varepsilon &
    \stepto[]{\statevar}{\statevar[']}{\event} &
    \stepton{\statevar[']}{\statevar['']}{\trace}
  }{
    \stepton[n+1]{\statevar}{\statevar['']}{\event\cdot\trace}
  }{n-step-trans}
  %
  \typerule{n-step-ign}{
    \stepto[]{\statevar}{\statevar[']}{\varepsilon} &
    \stepton{\statevar[']}{\statevar['']}{\trace}
  }{
    \stepton[n+1]{\statevar}{\statevar['']}{\trace}
  }{n-step-ign}

  \caption{Star-step and $n$-step relation for a given step relation $\stepto{\statevar[_{1}]}{\statevar[_{2}]}{\event}$}\label{fig:starstep:and:nstep}
\end{figure}
%
The \emph{behavior} of a whole program $w$ ($\behav{w}$) is a set containing all traces the program may emit during execution.
Note that for deterministic languages, $\behav{w}$ is a singleton set for any $w$.
With this, we can define satisfaction:

\begin{definition}[Property Satisfaction]\label{def:propsat}
  A \emph{whole} program $w$ satisfies $\pi$ ($\sat{w}{\pi}$) iff $\behav{w}\in\lift{\pi}$.
\end{definition}
\begin{definition}[Class Satisfaction]\label{def:classsat}
  A \emph{whole} program $w$ satisfies $\class$ ($\sat{w}{\class}$) iff $\behav{w}\in\class$.
\end{definition}

\subsection{Attacker Model and Robust Trace-Property Preservation}\label{subsec:bg:rtp}

Partial programs cannot be executed, because they lack definitions for some symbols.
Providing definitions for symbols is usually the job of the \emph{linker}.
We denote linking two programs $\progvar[_{1}],\progvar[_{2}]$ as $\link{\progvar[_{1}]}{\progvar[_{2}]}$.
For sake of simplicity, we assume that linking does not fail.
With linking, we can define robust satisfaction analogously to \Cref{def:propsat,def:classsat}:

\begin{definition}[Robust Property Satisfaction]\label{def:proprsat}
  A program $\progvar$ robustly satisfies $\pi$ ($\rsat{\progvar}{\pi}$) iff for any program $\contextvar$ we have $\sat{\link{\contextvar}{\progvar}}{\pi}$. (similarily for $\rsat{\progvar}{\class}$)
\end{definition}

Robust satisfaction gives rise to a maximally powerful attacker model: Any other partial program, here referred to as context $\contextvar$, is the attacker.
So, for example, if the programming language of $\progvar$ and $\contextvar$ allows guessing of memory addresses, then $\contextvar$ is able to read and write any position in memory and therefore guess the location of secrets.

We expect compilers to preserve robust satisfaction.
In this paper, compilers are modelled as function from one programming language's syntax to the other.
Since this paper considers composing compilers, we'll adopt notation similar to tombstone diagrams to easen the formal burden.
So, we denote a compiler from $\src{L}$ to $\trg{L}$ as $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and we refer to $\src{L}$ as \emph{source} language and to $\trg{L}$ as \emph{target} language.
Composing $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ with $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$ is denoted as $\cc{\src{L}\to\trg{L}\to\obj{L}}{\src{\bullet}}$ and means first running $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ and plugging the output into $\cc{\trg{L}\to\obj{L}}{\trg{\bullet}}$.
Contrary to tombstones, we don't consider a compiler's implementation language.
We now define the compiler-property robust trace-property preservation.

\begin{definition}[Robust Trace-Property Preservation]\label{def:rtp}
  A compiler $\cc{\src{L}\to\trg{L}}{\src{\bullet}}$ robustly preserves a class of properties $\class$ ($\rtp{\cc{\src{L}\to\trg{L}}{\src{\bullet}}}{\class}$) iff for any property $\pi\in\class$ and $\src{p}$ written in $\src{L}$, given $\rsat{\src{\progvar}}{\pi}$, then $\rsat{\cc{\src{L}\to\trg{L}}{\src{\progvar}}}{\pi}$.
\end{definition}

Note that in the conclusion of \Cref{def:rtp} there is a target-level context.
So, programs that robustly satisfy a property are hardened against a maximally malicious environment with respect to the power of the target language.

\section{Composing Properties}\label{sec:compprop}
\Cref{subsec:propdefs} introduces a ,,most general'' trace model and defines properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{cct}) that our case study (\Cref{sec:casestud}) on it.
By means of taking these properties as an example, \Cref{subsec:monitors} defines monitors for them and an agreement relation between a monitor's trace model with the most general trace model.
Then, we define monitor-based satisfaction and we prove that trace-property satisfaction is implied by it.
Since monitor-based satisfaction is easier to reason about, this makes our later proofs considerably simpler.

\subsection{Most General Trace-Model and Example Property Definitions}\label{subsec:propdefs}

\begin{definition}[\glsfirst{tms}]

\end{definition}
\begin{definition}[\glsfirst{sms}]

\end{definition}
\begin{definition}[\glsfirst{ms}]

\end{definition}
\begin{definition}[\glsfirst{scct}]

\end{definition}

\subsection{Monitors and Trace Agreements}\label{subsec:monitors}

\section{Composing Robustly Secure Compilers}\label{sec:compcomp}
This section studies different kinds of secure compiler compositions and their corresponding proofs.
Most notably, it looks at the sequential composition of secure compilers and shows a theorem that allows for modular proof engineering: compilers need not to robustly preserve the same property in order to be securely composed.
Moreover, it demonstrates that the order does not matter for composing compilers with the same source and target language.
